Networked Character Devices

### 2011-12-11 report/paper.tex
paper: client implementation section

@@

\subsubsection{udev}

>>>>

\lstinputlisting[
  caption={The NCD kernel module has a number of roles.},
  label=lst:client
]{cdev-driver.c}

The client resides entirely within a kernel module \footnote{The merits
of including everything within a kernel module are discussed in section
\ref{sec:currentstatus}.}. A typical userspace program can be sloppy
with cleanup because all memory is freed on termination. In contrast, a
kernel module must be very careful because its lifetime is shorter than
the kernel itself. Therefore it is very import to correctly handle
errors and cleanup all memory in the \texttt{module\_init} and
\texttt{module\_exit} routines.

Some of the basic features of the NCD kernel module are listed in
listing \ref{lst:client}. Firstly, NCD dynamically requests a major and
minor number from the running kernel \footnote{Statically assigned
major/minor numbers are no longer required and are discouraged for new
modules.}. These numbers are stored in a \texttt{struct dev\_t}. Once
these numbers are assigned, the kernel will pass all file operations on
character devices with the NCD major/minor number to the NCD module.

The \texttt{struct class} is somewhat synonymous with a kernel
subsystem. Because the major/minor number is dynamic, it is with this
class that userspace is able to persistently refer to NCD character
device nodes (see \ref{subsec:udev}).

\texttt{Struct file\_operations} is used to define the handlers for all
file operations which we support for our character devices \footnote{See
\texttt{linux/fs.h} for a complete list of file operations.}.

Most importantly, \texttt{struct cdev} ties together the major/minor
number with the given set of file operations. With this it is possible
for a kernel module to register for many major/minor numbers (typically
a contiguous range), while having fine-grained control over which
subsets are handled by which distinct groups of file operations. This
will be important in the future as specific subsystems are handled
differently by NCD to increase functionality.

Finally, \texttt{struct device} is used to actually create a character
device node under \texttt{/dev} in userspace. Unfortunately, not
everything we might want to do to this device can be accomplished in the
kernel (permissions for example).

@@

\subsubsection{udev}
\label{subsec:udev}
